---
title: "Template for a Base Case"
format: html
editor: visual
---

## Design

We consider a randomized phase III study.

-   Start with one placebo arm and a dose arm, with a 2:1 randomization ratio.
-   Planned trial duration is up to 40 months.
-   Randomize 50 patients per month for the first 6 months, then 65 patients per month until 650 patients are recruited.
-   Dropout rate is 5% at month 12 and 12% at month 24.
-   Two time-to-event endpoints `PFS` and `OS`.
    -   Medians of `PFS` in placebo and treated arms are 8.3 and 11.5 months, approximately a hazard ratio 0.72
    -   Medians of `OS` in placebo and treated arms are 14.5 and 19.3 months, approximately a hazard ratio 0.75
-   `PFS` and `OS` are both tested with one-sided `logrank` test, with the overall family-wise error rate controlled at $\alpha = 0.025$ under a graphical testing procedure.
    -   $\alpha$ is split between `PFS` (0.01) and `OS` (0.015).
-   A group sequential design is adopted
    -   Interim analysis is performed for both `PFS` and `OS` when 250 `PFS` events are observed.
    -   No futility analysis.
    -   Final analysis is performed for both endpoints when all planned patients are randomized and at least 400 `OS` events are observed. However, if the target number of `OS` events cannot be reached, the final analysis is performed at month 40.
    -   For `PFS`, Lan & DeMets $\alpha$-spending is used as an approximation to the O’Brien & Fleming boundaries.
    -   For `OS`, the Pocock boundary is adopted.

## Implementation

### Loading relevant packages

```{r}
#| output: false
library(dplyr)
library(TrialSimulator)
library(kableExtra)
set.seed(12345)
```

### Helper functions

In this section, we define helper functions that are task-specific. For example, statistician may simulate correlated `PFS` and `OS` in various ways. For the purpose of illustration, we simulate independent endpoints, i.e. `corr` is currently a dummy argument.

```{r}
#' Simulate correlated PFS and OS. Marginally, both endpoints follow exponential distribution. 
#' @param n integer. Number of samples. 
#' @param median_pfs numeric. Median of PFS in an arm. 
#' @param median_os numeric. Median of OS in an arm. 
#' @param corr numeric. Parameter of correlation between PFS and OS. 
correlated_pfs_os <- function(n, median_pfs, median_os, corr = .0){
  pfs <- rexp(n, rate = log(2) / median_pfs)
  os <- rexp(n, rate = log(2) / median_os)
  
  data.frame(pfs = pfs, os = os, pfs_event = 1, os_event = 1)
}
```

### Define Endpoints and Arms

::: panel-tabset
## Placebo

```{r}
pfs_and_os_in_pbo <- endpoint(name = c('pfs', 'os'), type = c('tte', 'tte'), 
                              generator = correlated_pfs_os, 
                              median_pfs = 8.3, median_os = 14.5)
pbo <- arm(name = 'placebo')
pbo$add_endpoints(pfs_and_os_in_pbo)
```

## Treated

```{r}
pfs_and_os_in_trt <- endpoint(name = c('pfs', 'os'), type = c('tte', 'tte'), 
                              generator = correlated_pfs_os, 
                              median_pfs = 11.5, median_os = 19.3)
trt <- arm(name = 'treated')
trt$add_endpoints(pfs_and_os_in_trt)
```
:::

### Define A Trial

We use the Weibull distribution to model dropout. The `TrialSimulator` package provides a helper function to compute the shape and scale parameters of the Weibull distribution.

```{r}
weibullDropout(time = c(12, 24), dropout_rate = c(.05, .12))
```

To define arms and a trial, we can

```{r}
#| output: false
accrual_rate <- data.frame(end_time = c(6, Inf), 
                           piecewise_rate = c(50, 65))

trial <- trial(
  name = 'trial-123', n_patients = 650, duration = 40, 
  enroller = StaggeredRecruiter, accrual_rate = accrual_rate, 
  dropout = rweibull, shape = 1.32, scale = 114.4
)

trial$add_arms(sample_ratio = c(1, 2), pbo, trt)
```

### Define Trial Milestones and Actions

Since the only adaptation for an endpoint is via group sequential design, we can use `logrank` p-values of cumulative data, and independent increment is not adopted.

::: panel-tabset
## Interim

```{r}
action_at_interim <- function(trial, milestone_name){
  
  locked_data <- trial$get_locked_data(milestone_name)
  
  #' codes below this line are customized to conduct statistical tests and save results
  
  pfs <- fitLogrank(endpoint = 'pfs', placebo = 'placebo', data = locked_data)
  
  os <- fitLogrank(endpoint = 'os', placebo = 'placebo', data = locked_data)
  
  trial$bind(value = pfs, name = 'pfs_result')
  trial$bind(value = os, name = 'os_result')
}
```

## Final

```{r}
#' action at final analysis is very similar to the one at interim, 
#' except that a graphical test is performed
action_at_final <- function(trial, milestone_name){
  
  locked_data <- trial$get_locked_data(milestone_name)
  
  #' codes below this line are customized to conduct statistical tests and save results
  
  pfs <- fitLogrank(endpoint = 'pfs', placebo = 'placebo', data = locked_data)
  
  os <- fitLogrank(endpoint = 'os', placebo = 'placebo', data = locked_data)
  
  trial$bind(value = pfs, name = 'pfs_result')
  trial$bind(value = os, name = 'os_result')
  
  #' extract logrank test results
  pfs_result <- trial$get_custom_data('pfs_result')
  os_result <- trial$get_custom_data('os_result')
  
  #' graphical test
  gt <- 
    GraphicalTesting$new(
      alpha = c(0.01, 0.015), 
      transition = matrix(c(0, 1, 1, 0), nrow = 2), 
      alpha_spending = c('asOF', 'asP'), 
      planned_max_info = c(530, 400), 
      hypotheses = c('PFS', 'OS'), 
      silent = TRUE
    )
  
  #' input for graphical test
  stats <- rbind(pfs_result, os_result) %>% 
    mutate(order = c(1, 2, 1, 2)) %>% 
    mutate(hypotheses = c('PFS', 'PFS', 'OS', 'OS')) %>% 
    mutate(is_final = c(FALSE, TRUE, FALSE, TRUE)) %>% 
    mutate(max_info = c(530, 530, 400, 400))
  
  print(stats)
  
  gt$test(stats)
  
  test_result <- gt$get_current_testing_results()
  
  milestone_time <- c(trial$get_milestone_time('interim analysis'), 
                      trial$get_milestone_time('final analysis'))
  
  
  #' save testing decisions for PFS and OS
  trial$save(value = test_result$decision[test_result$hypothesis == 'PFS'], 
             name = 'pfs_decision')
  
  trial$save(value = test_result$decision[test_result$hypothesis == 'OS'], 
             name = 'os_decision')
  
  #' save time of making the decisions for PFS and OS
  trial$save(value = milestone_time[test_result$order[test_result == 'PFS']], 
             name = 'pfs_decision_time')
  
  trial$save(value = milestone_time[test_result$order[test_result == 'OS']], 
             name = 'os_decision_time')
  
  
}
```
:::

We plan two trial milestones for interim and final analysis.

::: panel-tabset
## Interim

```{r}
interim <- milestone(name = 'interim analysis', 
                     trigger_condition = 
                       eventNumber(endpoint = 'pfs', n = 250), 
                     action = action_at_interim)
```

## Final

```{r}
final <- milestone(name = 'final analysis', 
                   trigger_condition = (
                     enrollment(n = 650) & 
                       eventNumber(endpoint = 'os', n = 400)
                     ) | calendarTime(time = 40), 
                   action = action_at_final)
```
:::

We then register the two trial milestones to a listener who will monitor the trial.

```{r}
#| output: false
listener <- listener()
#' register trial milestones with the listener
listener$add_milestones(interim, final)
```

### Execute a Trial

To execute a trial, we call the member function `run` in the controller of the trial and listener. Here we simulate one trial, so that a plot of cumulative events are generated automatically for each of the endpoints. We can set `n` to a larger integer for more replicates.

```{r}
controller <- controller(trial, listener)
controller$run(n = 1, plot_event = TRUE)
```

### Extract the Output

We can display information saved for each of the simulated trials by the member function `get_output`.

```{r}
controller$get_output() %>% 
  kable(escape = TRUE) %>% 
  kable_styling(bootstrap_options = "striped", 
                full_width = FALSE,
                position = "left") %>%
  scroll_box(width = "100%")
```
