---
title: "Simulation of Group Sequential Trial with a Time-to-Event Endpoint"
author: "Han Zhang"
date: "`r Sys.Date()`"
output: 
  html_document: 
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(class.source = "foldable")
```

```{r aljai, message=FALSE, echo=FALSE}
library(targets)
library(rpact)
library(TrialSimulator)
library(data.table)
library(kableExtra)
library(tidyverse)
```

This document replicates simulation result of a group sequential design for a time-to-event endpoint. Analytic calculation is based on `rpact::getPowerSurvival`, and simulation is based on 10,000 replicates using `TrialSimulator`. 

The design is under the following assumptions: 

- Group sequential with one interim at information fraction 66% using an O'Brien-Fleming type $\alpha$-spending function, one-sided level 2.5%, and power 80%. 

- Exponential distributed `PFS` with a median of 60 months in placebo. 

- Hazard ratio 0.74 between treatment and placebo arms. 

- Dropout rate 2.5% at 12 months in both arms. 

- Linear ramp up recruitment rate until the 6th month at 42 patients per month. 

- 1200 patients are 1:1 randomized into two arms. 

This example is discussed in `rpact` [vignette](https://www.rpact.org/vignettes/planning/rpact_survival_simulation_examples/). 

## Analytic Calculation Using `rpact`

```{r algna}
design <- getDesignGroupSequential(
    informationRates = c(0.66, 1),
    typeOfDesign = "asOF", 
    sided = 1, 
    alpha = 0.025, 
    beta = 1 - 0.8
)

sampleSizeResult <- getSampleSizeSurvival(
  design,
  median2 = 60, 
  hazardRatio = 0.74,
  dropoutRate1 = 0.025, 
  dropoutRate2 = 0.025, 
  dropoutTime = 12,
  accrualTime = c(0, 1, 2, 3, 4, 5, 6),
  accrualIntensity = c(6, 12, 18, 24, 30, 36, 42),
  maxNumberOfSubjects = 1200
)

sampleSizeResult |> summary()
```

## Simulation Using `TrialSimulator`

```{r laeig}
tar_load(example1)

with(example1, 
     data.table(
       'Interim analysis time' = mean(`event_time_<interim analysis>`), 
       'Final analysis time' = mean(`event_time_<final analysis>`), 
       'Power at interim' = mean(`decision_<interim>` == 'reject'), 
       'Overall power' = mean(`decision_<interim>` == 'reject' | 
                                `decision_<final>` == 'reject'), 
       'Expected Number of enrolled patients' = mean(ifelse(`decision_<interim>` == 'reject', 
                                                            `n_events_<interim analysis>_<patient_id>`,  
                                                            `n_events_<final analysis>_<patient_id>`)), 
       'Expected trial duration' = mean(ifelse(`decision_<interim>` == 'reject', 
                                               `event_time_<interim analysis>`, 
                                               `event_time_<final analysis>`))
     ) %>% 
       kable()
)

```

