---
title: "Simulation of Group Sequential Trial with a Binary Endpoint"
author: "Han Zhang"
date: "`r Sys.Date()`"
output: 
  html_document: 
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(class.source = "foldable")
```

```{r aljai, message=FALSE, echo=FALSE}
library(targets)
library(rpact)
library(TrialSimulator)
library(data.table)
library(kableExtra)
library(tidyverse)
```

This document replicates simulation result of a group sequential design for a time-to-event endpoint. Analytic calculation is based on `rpact::getPowerSurvival`, and simulation is based on 10,000 replicates using `TrialSimulator`. 

The design is under the following assumptions: 

- Group sequential with two interim and one final analysis at information fraction 50%, 75% and 100% using an O'Brien-Fleming type $\alpha$-spending function, one-sided level 2.5%, and power 90%. 

- In total, 266 patients are 1:1 randomized into two arms. Thus, 133 and 200 patients at the first and secondar interim. 

- Failure rates are $\pi_1 = 0.4$ in treatment and $\pi_2 = 0.6$ in control. 

- No dropout and readout is 0 as `rpact::getSampleSizeRates` does not support these settings. 

- Recruitment rate is set for `TrialSimulator` only as `rpact::getSampleSizeRates` does not evaluate interim time. 

- Non-binding futility bounds are 0 and 0.5 for the first and second interim on normal approximated z-value. In `TrialSimulator`, we calculate the z-value of the Farrington-Manning test of rate difference. 

A similar example is discussed in `rpact` [vignette](https://www.rpact.org/vignettes/planning/rpact_binary_planning_example/). 

## Analytic Calculation Using `rpact`

```{r algnaa}
design <- getDesignGroupSequential(
    informationRates = c(0.5, 0.75, 1), alpha = 0.025, beta = 1 - 0.9,
    futilityBounds = c(0, 0.5), 
    typeOfDesign = 'asOF'
)
r <- getSampleSizeRates(design, pi1 = 0.4, pi2 = 0.6)
r |> summary()
```

## Simulation using `rpact`

```{r aoeiega}
getSimulationRates(
  design, 
  pi1 = 0.4, pi2 = 0.6, 
  plannedSubjects = c(133, 200, 266), 
  directionUpper = FALSE, 
  maxNumberOfIterations = 10000
) |> summary()
```

## Simulation Using `TrialSimulator`

```{r laeig}
tar_load(example2)

example2 <- example2 %>%
  head(10000) %>% 
  mutate(
    `Number of enrolled patients` = 
      case_when(
        `decision_<interim1>` == 'reject' ~ .data[['n_events_<interim 1>_<patient_id>']], 
        `decision_<interim2>` == 'reject' ~ .data[['n_events_<interim 2>_<patient_id>']], 
        TRUE ~ .data[['n_events_<final>_<patient_id>']]
      )
  )

with(example2, 
     data.table(
       'Power at 1st interim' = mean(`decision_<interim1>` == 'reject'), 
       'Power at 2nd interim' = mean(`decision_<interim2>` == 'reject'), 
       'Overall power' = mean(`decision_<interim1>` == 'reject' | 
                                `decision_<interim2>` == 'reject' | 
                                `decision_<final>` == 'reject'), 
       'Expected number of enrolled patients' = mean(`Number of enrolled patients`), 
       'Exit Prob at 1st interim' = mean(`decision_<interim1>` == 'reject'),
       'Exit Prob at 2nd interim' = mean(`decision_<interim1>` == 'accept' & 
                                           `decision_<interim2>` == 'reject'), 
       'Exit Prob at 1st Futility' = mean(`futility_<interim1>`), 
       'Exit Prob at 2nd Futility' = mean(`futility_<interim2>`)
     ) %>% 
       kable()
)

```



