---
title: "Power of Group Sequential Trial with a Time-to-Event Endpoint"
author: "Han Zhang"
date: "`r Sys.Date()`"
output: 
  html_document: 
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(class.source = "foldable")
```

```{r aljai, message=FALSE, echo=FALSE}
library(targets)
library(rpact)
library(TrialSimulator)
library(data.table)
library(kableExtra)
library(tidyverse)
```

This document replicates simulation result of a group sequential design for a time-to-event endpoint. Analytic calculation is based on `rpact::getPowerSurvival`, and simulation is based on 10,000 replicates using `TrialSimulator`. 

The design is under the following assumptions: 

- Group sequential with three interim at information fraction 60%, 70% and 80% using an O'Brien-Fleming type $\alpha$-spending function, one-sided level 2.5%. 

- Exponential distributed `PFS` with an event rate 20% at 12 months in placebo. 

- Hazard ratio 0.85 between treatment and placebo arms. 

- Dropout rate 10% at 12 months in both arms. 

- Linear ramp up recruitment rate until the 6th month at 42 patients per month. 

- 3000 patients are 1:1 randomized into two arms. 

- Trial is ended when 1000 PFS events are observed. 

## Analytic Calculation Using `rpact`

```{r algna}

design <- getDesignGroupSequential(
  informationRates = c(.6, .7, .8, 1.0),
  typeOfDesign = 'asOF', 
  alpha = .025, 
  sided = 1
)

getPowerSurvival(
  design, 
  thetaH0 = 1.0, 
  pi2 = .2, 
  eventTime = 12, 
  directionUpper = FALSE,
  hazardRatio = .85, 
  dropoutRate1 = .1, 
  dropoutRate2 = .1, 
  dropoutTime = 12, 
  maxNumberOfSubjects = 3000, 
  maxNumberOfEvents = 1000 
) |> summary()
```

## Simulation Using `TrialSimulator`

```{r laeig}
tar_load(example3)

with(example3 %>% head(10000), 
     data.table(
       #'1st interim analysis time' = mean(`event_time_<interim 1>`), 
       #'2nd interim analysis time' = mean(`event_time_<interim 2>`), 
       #'3rd interim analysis time' = mean(`event_time_<interim 3>`), 
       #'Final analysis time' = mean(`event_time_<final>`), 
       'Power at 1st interim' = mean(`decision1` == 'reject'), 
       'Power at 2nd interim' = mean(`decision2` == 'reject'), 
       'Power at 3rd interim' = mean(`decision3` == 'reject'), 
       'Power at final' = mean(`decision4` == 'reject'),
       'Expected Number of PFS events' = mean(ifelse(`decision1` == 'reject', 
                                                            `n_events_<interim 1>_<pfs>`,  
                                                            ifelse(`decision2` == 'reject', 
                                                                   `n_events_<interim 2>_<pfs>`, 
                                                                   ifelse(`decision3` == 'reject', 
                                                                          `n_events_<interim 3>_<pfs>`, 
                                                                          `n_events_<final>_<pfs>`))))
     ) %>% 
       kable()
)

```

